Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-08 chatbot_classification/aaai_2021/paper.tex
version 04.   modified:   paper.tex   new file:   paper.ver04.pdf

  {Add joint NER and classification problem here.}

@@

  In  this  section,  we  conduct  extensive  experiments  over  5 task-specific
  datasets  to  evaluate  the  performance  of  our  proposed SFC system and the
  effectiveness  of  the  methodology  of  multi-task  learning  and
  joint-model structure under low-data settings.

  \subsection{Datasets}
  The  datasets  used in our experiments aims to help us study low-resource text
  classification  problem  under  task-specific conversational chatbot scenario.
  Therefore,  we did experiments on datasets under low-resource chatbot corpus setting,
  which  is a large number of class labels (50-250) with a few examples(5-20) in
  short  text  for  each  class, by sampling smaller subsets from the full scale
  when the raw datasets are oversized, see Table \ref{tbe:dataset statistic}. Besides, we also did extensive experiments with larger scale datasets with 30-50 examples per class to test the stability of our proposed SFC system under normal-size data settings, also see Table \ref{tbe:dataset statistic}.

  According to the task and domain, we categorize our experimental datasets into
  the following 3 categories:

  \textbf{Real-world task-specific faq chatbot dataset}

  \textbf{ITG}  is  a  real-world  FAQ  chatbot dataset composed of question and
  answer pairs about online English teaching. It contains 3,938 sample questions
  for  228  class labels, where each class label corresponds to an unique answer
  for a bunch of similar sample questions.

  We  divide the whole dataset into 3 folds randomly with approximately the same
  amount  of data within each fold, and then choose 2 folds as the training set,
  the  remaing  1  fold  as  the  test  set for our experiments. We can obtain 3
  different  combinations  among  the  3  folds,  and  we  will  use the average
  experimental  results  of  these 3 combinations as the final evaluation result
  for this dataset.

  \textbf{Customer review dataset for extreme classification}

  \textbf{Amazon-670K}   is   a   customer   product  review  dataset  for  text
  classification  task from the extreme classification repository with full data
  containing  670,091  class  labels,  285,176  training samples and 150875 test
  samples \cite{bhatia2016extreme}.  As  the  dataset  contains  multiple class
  labels  per  training sample, we keep only the first class label for each data
  point. Besides, we also randomly sampled 250 class labels from all the classes
  that  contain only 5-15 data points. In this way, we simulate the low-resource
  chatbot  corpus setting with the subset of 1,942 training samples and 716 test
  samples for 250 class labels.

  \textbf{Recently-released intention detection datasets}.

  From our point of view, task-specific chatbot usually starts with small amount
  of  data at the initial stage. As more and more customers start to communicate
  with  the  chatbot,  more data can be extracted from conversation log. Base on
  these  assumptions,  we  do  experiments  on  the following 3 intention detection
  dataset    under    sampling    method    similar    to    the   approach   in
  \cite{casanueva2020efficient}  but  in  more  sophisticated few-shot settings,
  while  keeping the same test sets for each experiment to see the change of the
  improvement scale with decreasing data amount.

@@

  \textbf{HWU64}   is  a  intention  detection  dataset  designed  for  home  robot
  scenario \cite{liu2019benchmarking}.  It aims at the specific task of capturing
  the  intention  for  different  user  requests  to  home  robot  and  finding the
  corresponding answer. The raw dataset contains 25,716 data points for 64 class
  labels  through  crowdsourcing,  which  raises  much  more  data  than  common
  real-world  circumstances. Therefore, we experiment with the setups where only
  5,  10,  15, 20, 30 and 50 training data points are available for each class label by
  random sampling.

  \textbf{CLINC150}  is a dataset designed for task-oriented systems with 23,700
  queries  that  are  short  and unstructured for 150 intents, in the same style
  made  by real users through crowdsourcing \cite{larson2019evaluation}. Here, we
  sample 5, 10, 15, 20, 30 and 50 training data points for each intention class label by random
  sampling for our experimental setup.

  \textbf{BANKING77}  is  a intention detection dataset for bank customer services.
  The    raw    dataset    contains    13,084   data   points   for   77   class
  labels \cite{casanueva2020efficient}.  Here,  we  sample 5, 10, 15, 20, 30, and 50 training
  data  points  for  each  class  label  by random sampling for our experimental
  setup.

  \subsection{Baselines}
  We choose 3 groups of typical baselines:

  \textbf{Group  1:  single-task  classification  model.} This group of baseline
  models  are  based  on  the network structure for classification model. As for
  non-pretrained  neural  network, we choose TextCNN \cite{kim2014convolutional}
  as  a representative baseline. As for the pretrained model, we choose RoBERTa,
  BERT  and  ALBERT,  all of which are among the most powerful pretrained models
  that  achieve  outstanding  results  in  many  NLP  tasks, in our experimental
  setting.  Since  we  apply  RoBERTa  in  our  SFC  system,  we naturally adopt
  single-task  classification  model  based on RoBERTa-large and RoBERTa-base as
  our  baseline  models.  Besides,  we also adopt the recently released baseline
  pre-trained model ALBERT-xxlarge \cite{lan2019albert} for comparison. Since we
  use few-shot settings different from that of \cite{casanueva2020efficient} for
  those  intention  detection  datasets,  we  also  use  BERT-large tuned on our
  few-shot settings as an essential reference baseline for experimental results.

  \textbf{Group  2: single-task similarity model.} This group of baseline models
  are   based   on   network   structure   for   sentence pair similarity   model. We choose RoBERTa-large for the pretrained model part as an
  powerful baseline. But during the evaluation step, we use elastic search to do
  a  rough  filter  for  candidate  sentence  pairs  to  guarantee  a reasonable
  inference time for each user input query.

  \textbf{Group  3:  SFC  system with ablation study.} This group of baseline is
  based  on  the network structure of two-stage SFC. We
  conduct  ablation study by introducing single task SFC system with only Task 1
  or  Task  2  to  investigate  the influence of each component within the total
  loss.

  \subsection{Implementation Details}
  For  our  SFC  system,  we  use RoBERTa-large as the pretrained model for both
  classification  task  and  sentence-pair similarity task. We further fine-tune
  our  SFC  system  on  task-specific  datasets with max sequence length of 128,
  learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
  of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.
  Also,  we  set  the  max number of the epoch to 100 and save the best model on
  validation  set based on test result in F-score. We also report the F-score as
  the main evaluation measure for all experimental results.

  In  two-stage  SFC, in equ. \ref{eq:two-stage_SFC_loss} the  weights of two tasks are set as $\alpha_S=0.125$ and
  $\alpha_K=0.875$  in our setting. The reason for giving task 2 relatively more
  weights  is  that  according  to  the  ablation  experiments we did in Section
  \ref{sec:exp},  task  2  contributes more to the overall performance. In joint
  SFC,  in equ. \ref{eq:joint_SFC_loss}, the  weights  of three tasks are set as $\alpha_S=0.25$, $\alpha_K=0.25$
  and $\alpha_C=0.5$ in our setting.

@@

  Table  \ref{tbe:table2}  presents  the  results for each experimental data and
  model  set-up  in detail. The performance of each model is measured in F-score
  on the fixed test set for each dataset. The overall performance of SFC system
  is  better  than  all  the  competitive  single-model  baseline models on five
  experimental   datasets.  This  observation  gives  us  a  general  idea  that
  classification model and similarity model can actually provide each other with
  complementary information, and combining them together into a joint system can
  effectively  enhance the overall capability based on richer extracted semantic
  information.  Furthermore, we also studied  the  relative performance among all the variants of SFC on  different data settings to explore the benefits brought by multi-task training and the joint-model structure.
  With all the experimental results shown in Table \ref{tbe:table2},  we did the following more detailed analysis from several different
  aspect.

>>>>

  {todo: Add joint NER and classification problem here.}

@@

  \subsection{Datasets}
  Our experiments aim to study the short text classification in the low-resource
  environment  popular  in  the  task-specific conversional chatbot application.
  Hence we choose and configure set datasets to serve this study. Different from
  the  datasets from other short text classification research, such as semantics
  classification  with  very  limited  semantics labels and several sentences as
  input,   the   datasets  adopted  in  our  experiments  are  typically  having
  comparatively  large  number  of  class labels, ranging from several dozens to
  hundreds,  and  each  class  label  is  associated  with  a handful of labeled
  queries,  and  each  query  is  usually  one  sentence. Table \ref{tbe:dataset
  statistic} displays their statistics.

@@

  \begin{enumerate}
    \item \emph{ITG},  is  a proprietory FAQ dataset from real-world
    chatbot  project,  which is composed of question and answer pairs about online
    English teaching. It contains 3,938 sample questions for 228 class labels, and
    each class label corresponds to a unique answer.

    \item \emph{Amazon-670K}, is a customer product review dataset for
    text  classification  task  from  the  extreme  classification repository. The
    complete  dataset  contains 670,091 class labels, 285,176 training samples and
    150,875   testing   samples   \cite{bhatia2016extreme}.  As  each  sample  may
    correspond  to  multiple  class labels, we keep only the first one. We further
    filter  out the class labels that each one is associated with 5 to 15 samples,
    or  reviews,  to  mimic  the  chatbot scenario. From them, we sample 250 class
    labels as well as their samples to form a subset with 1,942 + 716 samples.

    \item \emph{HWU64}, is a intention detection dataset designed for
    home  robot  scenario \cite{liu2019benchmarking}. It aims at the specific task
    of  capturing  the  intention  for  different  user requests to home robot and
    finding  the corresponding answer. The raw dataset contains 25,716 data points
    for 64 class labels through crowdsourcing.

    \item \emph{CLINC150},  is  a dataset designed for task-oriented
    systems  with  23,700 queries that are short and unstructured for 150 intents,
    in    the    same   style   made   by   real   users   through   crowdsourcing
    \cite{larson2019evaluation}.

    \item \emph{BANKING77}, is a intention detection dataset for bank
    customer  services.  The  raw dataset contains 13,084 data points for 77 class
    labels \cite{casanueva2020efficient}.
  \end{enumerate}

  Regarding  the first two datasets, we conduct 3-fold cross validata experients
  that   70   percent   as  training,  15  percent  as  validation  and  testing
  respectively,  and  report  the  averaged  testing results. Regarding the last
  three datasets, we conduct experiments using a sampling method similar to that
  in   \cite{casanueva2020efficient}   yet  in  a  more  sophisticated  few-shot
  settings.  We  fix  a  test  data for each one, and examine their performances
  using  5,  10,  15,  20,  30,  and 50 samples per class label respectively for
  training.  This  is important, as in practice, a task-specific chatbot usually
  starts  with  merely  a  handful of labeled data available in the early stage.
  Besides,  in  the  active  learning framework, building an effective anxiliary
  system  with limited resource is also quite important to help developers label
  data more efficiently.

  \subsection{System description}
  We choose three kinds of system to construct our baselines.

  \begin{enumerate}
    \item \emph{Pretrained model based classification system}, which consists of
    BERT,  RoBERTa,  and  ALBERT  \cite{lan2019albert}  based.  These pretrained
    models  are  practically  proven  to  achieve  outstanding  performances  in
    classification task and other NLP tasks.

    \item   \emph{Non-pretrained   model  based  classification  system},  which
    consists  of  a popular CNN based TextCNN \cite{kim2014convolutional}, and a
    recently  proposed label-embedding based LEAM (todo: reference). Empircally,
    a  non-pretrained  model  based  system  are  not  performing  as  well as a
    pretrained  models  based,  if  both is applicable in some task. Our listing
    these  non-pretrained model based systems here is to provide a comprehensive
    performance comparison on our datasets.

    \item   \emph{Pre-trained   model   based   similarity   model}.  We  choose
    RoBERTa-large  as  our  fundermental  module as our SFC systems are based on
    RoBERTa  too, and RoBERTa is empircally more effective than other pretrained
    models  in  the  short  text  classification  task  in  our  experiments. In
    inference,  we  use  an  elastic search to find a set of potential candidate
    labels for the similarity model, to guarantee a reasonable running time.
  \end{enumerate}

  Our SFC systems include two-stage SFC with ablation, trained with two tasks
  sperately and both, and joint SFC with all three tasks.

  \subsection{Implementation Details}
  For  our  SFC  systems,  we  use RoBERTa-large as the pretrained model for both
  classification  task  and  sentence-pair similarity task. We fine-tune
  our  SFC  system  on  our datasets with max sequence length of 128,
  learning  rate  of 1.5e-5 for RoBERTa-large pretrained model and learning rate
  of  5e-4  for  task-specific  layer  with polynomial decay and Adam optimizer.
  Besides,  we  set  the  max number of the epoch to 100 and save the best model on
  validation  set based on testing result in weighted F-score (todo: reference). 
  We report the F-score as the main evaluation measure for all experiments.

  In two-stage SFC, in equ. \ref{eq:two-stage_SFC_loss} the weights of two tasks
  are  set  as  $\alpha_S=0.125$ and $\alpha_K=0.875$ in our setting. The reason
  for  giving  task  2 relatively more weights is that according to the ablation
  experiments  we  did  in Section \ref{sec:exp}, task 2 contributes more to the
  overall  performance.  In  joint  SFC,  in  equ.  \ref{eq:joint_SFC_loss}, the
  weights  of  three  tasks  are  set  as  $\alpha_S=0.25$,  $\alpha_K=0.25$ and
  $\alpha_C=0.5$  in  our setting. Overrall, the task weights are less sensitive
  in Joint SFC than in two-stage SFC.

@@

  Table  \ref{tbe:table2}  presents  all results on the five datasets 
  model  set-up  in detail. The performance of each model is measured in F-score
  on the fixed test set for each dataset. 
  Overall,  two-stage SFC systems show moderate improvements over all  baseline
  systems.  This  observation  gives  us  a  general  idea  that
  classification model and similarity model can actually provide complementary information
  to each other, and combining them together can enhance the overall capability 
  based on richer extracted semantic information.  
  Furthermore, we also studied  the  performance among all the variants of SFCs on  
  different data settings to explore the benefits brought by multi-task training 
  and the joint-model structure.
